Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations

### 2024-03-28 Delaneyetal2024.tex
push to overleaf

\correspondingauthor{Ian Delaney}{ianarburua.delaney@unil.ch}
\ian{quick check on shear stress units: seems like Pa, not Pa/m}
\begin{keypoints}
\item Short water discharge fluctuations in subglacial channels are accommodated by water velocity, not channel size, as in subaerial systems.
\item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
\item Significant subglacial sediment transport capacity can persist across a range of water discharges, impacting the  timing of peak events.
\end{keypoints}

\begin{abstract}
  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with the water velocity and channel width.
  In subaerial channels, water discharge variations are accommodated largely by flow depth and width changes.
  In subglacial channels rather, water discharge changes primarily affect water velocity due to evolving channel geometry over hours to weeks.
  Here, hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
  Results also show that high subglacial sediment transport rates and water velocities can occur across wide-ranging water discharges.
  These findings are discussed in the context of understanding sediment discharge from glaciers with different hydrological forcings and assumptions of channel type.
\end{abstract}

@@

\section{Methods}
\label{sect:meth}
The models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \cite<Figure \ref{fig:cartoon}; >{shields1936}.
The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.

@@

To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the velocity of the flowing water.
To do this, we use a lumped hydraulics model from \citeA{werder2010b}, itself based upon  \cite{clarke1996}.

@@

% To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is sufficiently wide compared to its depth so that the hydraulic radius is well approximated by the flow depth,
the shear stress $\tau_f$ at the river bed is
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_f^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_f} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\nabla z_c$ is the channel slope, and $f_f$ is the friction factor for subaerial channels \cite{tucker1997}.
Channel width $w_f$ is
\begin{equation}
  \label{eq:wcf}
  w_f = k \, Q^{\alpha},
\end{equation}
%
where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
Water velocity, $v_f$, is given as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_f\,\rho_w}}.
\end{equation}
%
As above, the width-integrated shear stress is $w_f\tau_f$.

Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
Thus the channel size in the subaerial model has no history dependence on the discharge $Q_w$, where as the subglacial one does.

@@

The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}).
The subglacial and subaerial models represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
\icesheet{}  is exemplified by Leverett glacier's  thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
%However, both scenarios share a discharge with a similar, relative seasonal amplitude.

In the first experiment, for each glacier, the models are  applied to a reference test case  that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$%, i.e. a segment of a quarter-circle,
(Table~\ref{table:vpm}).
Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with the maximum observed water discharge for the first $4$ days of the study until there is no change in channel area. 
The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ (Table~\ref{table:vpm}).

The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
For brevity, these values together are referred to as ``model outputs'' in the text.
Model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which varies with the ordering of values.
Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.

@@

The effect of channel shape and roughness is established by running the model with random parameter values of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for values).
Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$, following measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
Additionally, to be accepted, model runs can only experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig3.png}
  \caption{
    Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
    Variables on the x-axis have been normalized to mean values.
    $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
    Plots are shown with discharge and outputs at $15$\unit{min} intervals.
  }
  \label{fig:Qw_vari}
\end{figure}

Examining the relationship between water discharge and the model outputs, a different relationship emerges between the subaerial and subglacial cases, due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari}).
For instance, in subglacial channels in \alpine{}, very high seasonal water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over ($17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, Figure~\ref{fig:Qw_vari}\,a).
In \icesheet{}, water velocities close to the seasonal mean value over the season can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).

@@

\begin{figure}[h]
  \centering
    \includegraphics[width=0.9\linewidth]{Fig4.png}
    \caption{Standard deviation of model outputs for different time aggregations with different subglacial and subaerial channel shapes and friction factors.
      Shaded areas denote the range of results from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
      Solid lines denote  mean values.
      Markers show aggregation periods ($15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
       }
    \label{fig:multi_run}
  \end{figure}

@@

Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial  decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, longer than which seasonal variations might be conflated.
Note that the timing of peak sediment transport conditions still occurs across these periods, and they are are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).

@@

Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counterparts \cite<Figures~\ref{fig:Qw_vari}~and~\ref{fig:multi_run}>{alley1997}.
In subglacial R\"othlisberger channels in steady state with water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
However, subglacial channels rarely operate in a steady state \cite<e.g.>{gimbert2016}.
For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $w_g\, \tau_{g}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
\ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike with the evolving channel size here where a disparate relationship with water discharge occurs (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}). 
Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than the subglacial channel's response of hours to days.

@@

Previous work suggests that this co-varying relationship between sediment discharge capacity in  subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge.

Furthermore, the greater variations in subglacial sediment transport capacity compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization begins, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
Sediment exhaustion through this process may  explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.

@@

Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting the ability of subglacial  sediment transport models to capture certain events.
The model outputs' variability in subglacial channels approaches that of subaerial ones over $1-5$ days of  aggregation (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
Note that although variability between the events could remain comparable, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity (Figure~\ref{fig:model_outs}).

@@

    \textbf{Parameters and Constants}  & & &\\
    Gravitational constant&$g$& $9.81$&$\mathrm{m\,s^{-2}}$\\
    Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
    Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
    Hooke angle of channel & $\beta$ & $\frac{\pi}{6}$ & \unit{rad}\\
         && $\beta$:($\frac{\pi}{10}$, $\pi$) & \\
    Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5$ or $16$, $3$ & $\mathrm{(-)}$ \\
         && $f_g,f_f$:($0.01$, $21$) & \\
    Glacier thickness &$h_{ice}$& $225$ or $740$  &\unit{m}\\
    Effective glacier thickness &$h_o$&$\frac{\rho_i}{\rho_w} h_{ice}$  &\unit{m}\\
    Effective glacier length &$l$& $7,000$ or $26,000$&\unit{m}\\
    Constant $1$ in Equation~\ref{eq:dS_dt} &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
    Constant $2$ in Equation~\ref{eq:dS_dt} &$C_2$&$3.7\times10^{-13}$&\unit{m}$^{-n}\,s^{-1}$\\
    Latent heat of fusion &$L$&$333.5 $&\unit{kJ\,kg}$^{-1}$\\
    Pressure melting coefficient &$c_t$&$7.5\times 10^{-8}$&\unit{K\,Pa}$^{-1}$\\
    Specific heat capacity of water &$c_p$&$4180$&\unit{J\,kg}$^{-1}$\unit{K}$^{-1}$\\
    Ice flow constant &$A$& $5.3\times10^{-24}$ &\unit{Pa}$^{-n}$\,$s^{-1}$\\
    Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\

    Gradient of channel bed (subaerial) &$\nabla z_c$ &$0.05$& $\mathrm{(-)}$\\
    Subaerial channel factor & $k$ &$3$ or $6.5$ & $\mathrm{s\,m^{-2}}$\\
     && $k$:($5$, $20$) & \\
    Channel geometry exponent &$\alpha$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
    \hline
  \end{tabular}
  \label{table:vpm}
\end{table}
\FloatBarrier

>>>>

\correspondingauthor{Ian Delaney}{ianarburua.delaney@unil.ch}

\begin{keypoints}
\item Short water discharge fluctuations in subglacial channels are accommodated by water velocity, not channel size, as in subaerial systems.
\item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
\item Significant subglacial sediment transport capacity can persist across a range of water discharges, impacting the  timing of peak events.
\end{keypoints}

\begin{abstract}
  Sediment transport capacity in both subaerial and subglacial channels depends on the shear stress exerted across the channel bottom, which varies with the water velocity and channel width.
  In subaerial channels, water discharge variations are accommodated  by flow depth and width changes, along with water velocity.
  In subglacial channels rather, water discharge changes primarily affect solely water velocity due to evolving channel geometry over hours to weeks.
  Here, hydrographs from an Alpine glacier and the Greenland Ice Sheet are used to drive subglacial and subaerial hydraulics models, revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
  Results also show that high subglacial sediment transport rates and water velocities can occur across wide-ranging water discharges.
  These findings are discussed in the context of understanding sediment discharge from glaciers with different hydrological forcings and assumptions of channel shape.
\end{abstract}

@@

Note that sediment discharge records are omitted from the analysis here given the role of sediment supply, in addition to capacity, on capturing variations in sediment discharge \cite<e.g.>{delaney2024}.

\section{Methods}
\label{sect:meth}
The models described below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
Both models use the measured discharge to calculate water velocity, shear stress, and width-integrated shear stress, upon which both suspended sediment and bedload transport depend \cite<Figure \ref{fig:cartoon}; >{shields1936}.
The evaluation of these variables omits the selection of a sediment transport relationship and a grain-size parameter \cite<e.g.>{shields1936,meyer1948}.

@@

To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel model accounts for the channel geometry and the velocity of the flowing water.
To do this, we use a lumped hydraulics model from \citeA{werder2010b}, itself based upon  \citeA{clarke1996}.

@@

% To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is sufficiently wide compared to its depth so that the hydraulic radius is well approximated by the flow depth,
the shear stress $\tau_f$ at the river bed is
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_f^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_f} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\nabla z_c$ is the channel slope, and $f_f$ is the friction factor for subaerial channels \cite{tucker1997}.
Channel width $w_f$ is
\begin{equation}
  \label{eq:wcf}
  w_f = k \, Q^{\alpha},
\end{equation}
%
where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
Also following the Darcy-Weisbach, subaerial water velocity, $v_f$, is given as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_f\,\rho_w}}.
\end{equation}
%
As above, the width-integrated shear stress is $w_f\tau_f$.

Note that this subaerial channel model is purely algebraic, whereas the subglacial model comprises a differential equation for the evolution of $S_g$.
Thus the channel size in the subaerial model has no history dependence on the discharge $Q_w$, where as the subglacial one does (Equation~\ref{eq:dS_dt}).

@@

The models above are applied to proglacial discharge records from the Fieschergletscher (scenario \alpine{}) and the Leverett Glacier (scenario \icesheet{}), using subglacial model outputs that are contrasted with subaerial ones.
Outputs of the models represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
To generalize these scenarios, \alpine{}  is exemplified by Fieschergletscher's relatively thin ice thickness \cite<$h_{ice}$: $225$\,\unit{m}>{grab2021}, low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
\icesheet{}  is exemplified by Leverett glacier's  thick ice  \cite<$h_{ice}$: $700$\,\unit{m}; >{morlighem2017}, high water discharge \cite<$\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$>{tedstone2017}  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
%However, both scenarios share a discharge with a similar, relative seasonal amplitude.

In the first experiment, the models are  applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{6}$%, i.e. a segment of a quarter-circle,
(Table~\ref{table:vpm}).
Both friction factors $f_r$ and $f_f$ are tuned so that reasonable water velocities \cite<$\sim\,1.6\,$\unit{m}\,\unit{s}$^{-1}$>{werder2010b,chandler2013} occur for both \alpine{} and \icesheet{}.
The initial condition of cross-sectional area $S_g$ is evaluated by applying a model spinup with the maximum observed water discharge for the first $4$ days of the study until there is no change in channel area. 
The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ (Table~\ref{table:vpm}).

The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
For brevity, these values together are referred to as ``model outputs'' in the text.
To quantify the link between sediment transport capacity and water discharge, model outputs are compared to water discharge from the glaciers using Spearman rank correlation, which varies with the ordering of values.
Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.

@@

The effect of channel shape and roughness is established by running the model with random parameter values of channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_f$; see Table~\ref{table:vpm} for value range).
Runs with parameter combinations are accepted if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}\,\unit{s}$^{-1}$ and $2$\,\unit{m}\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}\,\unit{s}$^{-1}$, following measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
Additionally, to be accepted, model runs can only experience a flotation fraction ($\frac{\Delta h}{2\,h_o}$) of above $1.2$ for less than $2.5$\,\% of the run.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig3.pdf}
  \caption{
    Relationship between water discharge and normalized velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet{} (d-f).
    Variables on the x-axis have been normalized to mean values.
    $R_g$ ($R_f$) shows the Spearman rank correlations for the subglacial (subaerial) outputs.
    Plots are shown with discharge and outputs at $15$\unit{min} intervals.
  }
  \label{fig:Qw_vari}
\end{figure}

Examining the relationship between water discharge and the model outputs, a different relationships emerge between the subaerial and subglacial cases, due to the hysteresis in channel size in the subglacial model (Equation~\ref{eq:dS_dt}).
Because there is no history dependence in  subaerial channels, each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress (Section~\ref{sect:sub_mode}).
This characteristic results in a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari}).

Conversely, the history dependence on channel size in subglacial channels means different outputs occur across a range of discharges.
For instance, in subglacial channels in \alpine{}, very high seasonal water velocity values and shear stresses can occur from a low water discharge  ($\sim\,4$\,\unit{m}$^3$\,\unit{s}$^{-1}$) to the maximum water discharge at over ($17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, Figure~\ref{fig:Qw_vari}\,a).
In \icesheet{}, water velocities close to the seasonal mean value can occur at water discharges between roughly $150$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $310$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
When considering the subglacial channel's evolving width, width-integrated shear stress  across the channel generally increases with water discharge, with improved rank correlation compared to water velocity or shear stress (\alpine{}, Figure~\ref{fig:Qw_vari} a--c).
Yet even the width-integrated shear stress  can vary substantially, with the highest values occurring at water discharge values ranging from roughly $11$ \,\unit{m}\,\unit{s}$^{-1}$ to over $17$ \,\unit{m}\,\unit{s}$^{-1}$.
The variability in width-integrated shear stress is less pronounced in the \icesheet{} scenario, where the hydrograph has less diurnal variability (Figure~\ref{fig:Qw_vari} c, f).

@@

\begin{figure}[h]
  \centering
    \includegraphics[width=0.9\linewidth]{Fig4.pdf}
    \caption{Standard deviation of model outputs for different time aggregations with different subglacial and subaerial channel shapes and friction factors.
      Shaded areas denote the range of results from the accepted $100$ parameter combinations (Section~\ref{sect:imp}).
      Solid lines denote  mean values.
      Markers show aggregation periods ($15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, and $15$\,\unit{d}).
       }
    \label{fig:multi_run}
  \end{figure}

@@

Results from  both \alpine{} and \icesheet{} scenarios show that the variability in subglacial  decreases with aggregations longer than approximately 1--5 days (Figure~\ref{fig:multi_run}).
Yet, substantial differences in the variability between the model outputs for subglacial and subaerial cases persist in periods of up to $15$ days, longer than which seasonal variations might be conflated.
Note that the timing of peak sediment transport conditions still occurs, across these periods, and they are are thus not captured in aggregating water discharge over longer periods (Figure~\ref{fig:model_outs}).

@@

Sediment transport capacity in subglacial channels responds more rapidly to changes in water discharge compared to their subaerial counterparts \cite<Figures~\ref{fig:Qw_vari}~and~\ref{fig:multi_run}>{alley1997}.
In subglacial R\"othlisberger channels in steady state with water discharge, the width-integrated shear stress response to water discharge is $ w_g\,\tau_{g}\, \propto\,  Q^{\sim\frac{4}{5}}$ (Table\,\ref{tab:eqs1}).
However, subglacial channels rarely operate in a steady state \cite<e.g.>{gimbert2016}.
For sufficiently short periods, channel width and wetted area are fixed in the subglacial system (akin to pipe flow) so that the response of width-integrated shear stress is $w_g\, \tau_{g}\, \propto\,  Q^{2}$ (Table\,\ref{tab:eqs1}).
\ref{sect:scaling} provides a complete description of sediment transport capacity in different drainage conditions.
Note that in pipe flow conditions, maximum sediment transport would coincide with peak water discharge and a co-varying relationship with water discharge would occur, unlike with the evolving channel size here where a disparate relationship with water discharge occurs (Section~\ref{sect:sub_mode}; Figure \ref{fig:model_outs} and \ref{fig:Qw_vari}). 
Similar adjustments in channel shape occur in subaerial channels \cite{phillips2016}, albeit at timescales longer than the subglacial channel's response time of hours to days.

@@

Previous work suggests that this co-varying relationship between sediment discharge capacity in  subaerial conditions could be noticeable $\sim20$\,\unit{km} downstream of the Leverett site, where a strong correlation persists between sediment plume size and the river's water discharge into the Kangerlussuaq fjord \cite{mcgrath2010}.
In contrast, in marine-terminating glacier catchments, a less consistent relationship may occur between water discharge or melt extent and sediment plume size \cite{tedstone2012}.
Here, ocean water pressurizes subglacial channels at the ice front \cite<e.g.>{how2017}, so the reduced correlation could result from the inconsistent relationship between subglacial sediment transport capacity and water discharge presented here.

Furthermore, the greater variations in subglacial sediment transport capacity compared to subaerial ones could cause a supply-limited regime to persist at many glaciers.
In subglacial systems, sediment's critical shear stress, or threshold at which sediment mobilization occurs, can be reached more frequently and across a range of water discharges, compared to subaerial systems (Figure~\ref{fig:Qw_vari}).
Sediment exhaustion through this process may  explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \cite{overeem2017}.

@@

Water discharge measurements at the hourly timescale remain unavailable in many catchments, further limiting the ability of subglacial  sediment transport models to capture certain events.
The model outputs' variability in subglacial channels approaches that of subaerial ones over $1-5$ days of  aggregation (Figure~\ref{fig:multi_run}), possibly resulting in parity in sediment transport capacity variations between subaerial and subglacial channels over these periods.
Note that although variability between the events could remain comparable, water discharge forcings aggregated over periods longer than days may omit key sediment discharge characteristics, such as the timing of peaks in sediment transport capacity from glaciers (Figure~\ref{fig:model_outs}).

@@

    \textbf{Parameters and Constants}  & & &\\
    Gravitational constant&$g$& $9.81$&$\mathrm{m\,s^{-2}}$\\
    Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
    Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
    Hooke angle of channel & $\beta$ & $\frac{\pi}{6}$ & \unit{rad}\\
         && ($\frac{\pi}{10}$, $\pi$) & \\
    Friction factor (subglacial, subaerial) & $f$, ($f_r$, $f_f$) &$5$ or $16$, $3$ & $\mathrm{(-)}$ \\
         && ($0.01$, $21$) & \\
    Glacier thickness &$h_{ice}$& $225$ or $740$  &\unit{m}\\
    Effective glacier thickness &$h_o$&$\frac{\rho_i}{\rho_w} h_{ice}$  &\unit{m}\\
    Effective glacier length &$l$& $7,000$ or $26,000$&\unit{m}\\
    Constant $1$ in Equation~\ref{eq:dS_dt} &$C_1$&$2.2\times10^{-5}$&\unit{m}$^{-1}$\\
    Constant $2$ in Equation~\ref{eq:dS_dt} &$C_2$&$3.7\times10^{-13}$&\unit{m}$^{-n}\,s^{-1}$\\
    Latent heat of fusion &$L$&$333.5 $&\unit{kJ\,kg}$^{-1}$\\
    Pressure melting coefficient &$c_t$&$7.5\times 10^{-8}$&\unit{K\,Pa}$^{-1}$\\
    Specific heat capacity of water &$c_p$&$4180$&\unit{J\,kg}$^{-1}$\unit{K}$^{-1}$\\
    Ice flow constant &$A$& $5.3\times10^{-24}$ &\unit{Pa}$^{-n}$\,$s^{-1}$\\
    Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\

    Gradient of channel bed (subaerial) &$\nabla z_c$ &$0.05$& $\mathrm{(-)}$\\
    Subaerial channel factor & $k$ &$3$ or $6.5$ & $\mathrm{s\,m^{-2}}$\\
     && ($5$, $20$) & \\
    Channel geometry exponent &$\alpha$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
    \hline
  \end{tabular}
  \label{table:vpm}
\end{table}
\FloatBarrier
